\Large \bf Vigor: Toward Carrier-Grade Software Network Functions

### 2016-12-05 doc/paper/paper.tex
Minor corrections.

This section motivates and explains at a high level what it's like to use Vigor
when developing NFs. For the sake of illustration, say we wish to develop a
trivial stateful NF that implements the discard protocol~\cite{rfc863}: an
infinite loop receives packets from one interface, discards the ones that target
port 9 (``discard port''), and forwards the others through another interface.
A possible implementation is on the Fig.~\ref{lst:orig}.

@@

Fig.~\ref{lst:vigor} shows the annotated code, ready for the verification with
Vigor. It is the result of applying the first two steps shown below.
\code{VIGOR_CHECK} (l.\ref{line:vigcheck}) adds additional property that we
would like to verify: ``we do not forward packets destined to port 9''.

@@

Vigor makes it possible for NF developers to verify their code without having to
understand verification. Note that stating invariants(step 3) for dataplane
software turns out to require no verification expertise.

@@

You can see one of the generated call sequences on Fig.~\ref{lst:call-sequence}.
It starts with \code{loop_invariant_produce} and continues then with calls to
\code{ring_full} and \code{ring_empty}, noting (with \code{assume}) that both
return \code{false}. They both return false in this particular trace, in other
traces they returned other values, but to define the context for the last
function call (l.\ref{line:tip-call}) we must repeat the history as closely as
possible. Vigor handles the result of the last call in the sequence very
differently. Instead of just telling the formal verification engine what
happened, it inserts an \code{assert}(l.\ref{line:assert}) reflecting the
symbolic model output to check that it is consistent with the postcondition of
the last function (Fig.~\ref{lst:contract}).

@@

Let us follow now the `port != 9' condition across the framework. During
exhaustive SE, SEE first learns the constraint in the \code{if} condition:
Fig.~\ref{lst:vigor} l.\ref{line:if-cond}. Here, SEE just records it in the call
sequence for model validation, and does not use further in the SE, because our
model of \code{ring_push_back} discards the packet being pushed. However, the
call to \code{ring_pop_front} (l.\ref{line:pop-front}) reintroduces this
constraint formulated in \code{packet_constraints}
(Fig.~\ref{lst:customizations}~l.\ref{line:c-pkt-cnstr}), according to the
symbolic model - Fig.~\ref{lst:symbolic-model}. Being applied to the fresh
packet from the ring, it satisfies the \code{VIGOR_CHECK}
(Fig.~\ref{lst:vigor}~l.\ref{line:vigcheck}).

During the operation, SEE produces call sequences that also capture the said
constraint. E.g. the line~\ref{line:assert} on Fig.~\ref{lst:call-sequence}
reflects the condition (Fig.~\ref{lst:customizations} l.\ref{line:c-pkt-cnstr})
that was added by the symbolic model (Fig.~\ref{lst:symbolic-model}
l.\ref{line:assume}).

@@

The component developer produces several model that capture the component
semantics in more or less details for different usecases. For a particular use
case --- a NF, a particular model may be inappropriate. An inappropriate model
may not compromise the verification result, because Vigor detects both
over- and under-approximations.

@@

Suppose now we had an underapproximate model: Fig.~\ref{lst:underapproximate}.
Then the last line in the call sequence check (Fig.\ref{lst:call-sequence}) in
this case would look different:
\begin{lstlisting}
/*@ assert(arg2.port == 0); @*/
\end{lstlisting}
Vigor model validation phase flags this inconsistency. Specifically, the
assertion fails, because the contract (Fig.~\ref{lst:contract}) specifies a
wider range for \code{arg2.port} than just \code{0}.

Different models may work well for different usecases. E.g. the overapproximate
model is just right for a buffer, that does not drop the packets with port 9.
The underapproximate model would be just right for a buffer that forwards only the
packets with port equal to 0.

>>>>

This section motivates and explains at a high level what it's like to use Vigor
when developing NFs. For the sake of illustration, say we wish to develop a
trivial stateful NF that implements the discard protocol~\cite{rfc863}: an
infinite loop receives packets from one interface, discards the ones that target
port 9 (``discard port''), and forwards the others through another interface.
A possible implementation is on Fig.~\ref{lst:orig}.

@@

Annotated code ready for verification with Vigor is shown on
Fig.~\ref{lst:vigor}. It is the result of applying the first two steps shown
below. \code{VIGOR_CHECK} (l.\ref{line:vigcheck}) adds an additional property
that we would like to verify: ``we do not forward packets destined to port 9''.

@@

Vigor makes it possible for NF developers to verify their code without having to
understand verification. Note that stating invariants (step 3) for dataplane
software turns out to require no verification expertise.

@@

You can see one of the generated call sequences on Fig.~\ref{lst:call-sequence}.
It starts with \code{loop_invariant_produce} and continues then with calls to
\code{ring_full} and \code{ring_empty}, noting (with \code{assume}) that both
return \code{false}. They both return false in this particular trace, in other
traces they return other values, but to define the context for the last function
call (l.\ref{line:tip-call}) we must repeat the history as closely as possible.
Vigor handles the result of the last call in the sequence very differently.
Instead of just telling the formal verification engine what happened, it inserts
an \code{assert}(l.\ref{line:assert}) reflecting the symbolic model output to
check that it is consistent with the postcondition of the last function
(Fig.~\ref{lst:contract}).

@@

Let us follow now the `port != 9' condition across the framework. During
exhaustive SE, SEE first learns the constraint in the \code{if} condition
(Fig.~\ref{lst:vigor} l.\ref{line:if-cond}). Here, SEE just records it in the
call sequence for model validation, and does not use it further in the SE,
because our model of \code{ring_push_back} discards the packet being pushed.
However, the call to \code{ring_pop_front} (l.\ref{line:pop-front}) reintroduces
this constraint formulated in \code{packet_constraints}
(Fig.~\ref{lst:customizations}~l.\ref{line:c-pkt-cnstr}), according to the
symbolic model - Fig.~\ref{lst:symbolic-model}. Being applied to the fresh
packet from the ring, it satisfies the \code{VIGOR_CHECK}
(Fig.~\ref{lst:vigor}~l.\ref{line:vigcheck}).

During the operation, SEE produces call sequences that also capture the said
constraint. For example, line~\ref{line:assert} on Fig.~\ref{lst:call-sequence}
reflects the condition (Fig.~\ref{lst:customizations} l.\ref{line:c-pkt-cnstr})
that was added by the symbolic model (Fig.~\ref{lst:symbolic-model}
l.\ref{line:assume}).

@@

The component developer produces several models that capture the component
semantics in more or less details for different use cases. For a particular use
case --- a NF, a particular model may be inappropriate. An inappropriate model
may not compromise the verification result, because Vigor detects both
over- and under-approximations.

@@

Suppose now we had an underapproximate model --- Fig.~\ref{lst:underapproximate}.
Then the last line in the call sequence check (Fig.\ref{lst:call-sequence}) in
this case would look different:
\begin{lstlisting}
/*@ assert(arg2.port == 0); @*/
\end{lstlisting}
Vigor model validation phase flags this inconsistency. Specifically, the
assertion fails, because the contract (Fig.~\ref{lst:contract}) specifies a
wider range for \code{arg2.port} than just \code{0}.

Different models may work well for different usecases. For instance the
overapproximate model is just right for a buffer, that forwards all the packets
(including the aimed to port 9). The underapproximate model would be just right
for a buffer that forwards only the packets with port equal to 0.
